Deciding to Disclose: Pregnancy and Alcohol Misuse

### 2014-07-16 Background.tex
Heuristics & descriptive

As noted, heuristic decision making stems from a contention that \citeauthor{Neumann1953} type rationality ignores the context of decision making.

@@

While normative theories are concerned with how decisions should be made, descriptive theories instead focus on how decisions are made.

>>>>

As noted, heuristic decision making stems from a contention that \citeauthor{Neumann1953} type rationality ignores the both context of decision making, and a lack of correspondence between predicted and actual human decisions (see, for example the Allais paradox \citep{Society2013}, and subsequent empirical support \citep{Oliver2003,Burke1996}). Arguably, this begins with \citet{Simon1956}, who suggested that humans do not attempt to make optimal choices, but to satisfice and choose the first `good enough' option. While noting that this will often achieve the same result, the claim is that humans exhibit bounded rationality \citep{Simon2000} arising from inherent limits to cognition.

\citet{Gigerenzer1996} take the concept of bounded rationality further, and argue for what they term \ac{FFH}. This recasts rationality as bound to the context of the behaviour - a rational approach to choosing the right mate might well require checking every possible partner, but given finite time, memory, and so on rapidly becomes nonsensical.

@@

While heuristic theories arguably fall under the purview of the descriptive, the wider tendency is towards what are in essence patches to normative models. The most influential models in this class derive from \ac{PT} \citep{Kahneman1979}, which combines a set of heuristics based on observed decision behaviour, with distortions to the perception of probability, and the value of outcomes. \citet{Tversky1992} subsequently addressed key issues present in their original formulation by introducing \ac{CPT}, which allows for non-binary decisions, at the expense of the heuristics. The essence then, is that high and low probabilities are treated differently, and the subjective value of a loss differs from the equivalent gain (losing your shirt is worse than winning a shirt is good).
